---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyr)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(data.table)
theme_set(theme_bw())
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))    
d = read.csv("ling245B/Manko245B_Project/analysis/01_experiment/data/SIS-trials.csv", header = TRUE)
sinfo = read.csv("ling245B/Manko245B_Project/analysis/01_experiment/data/SIS-subject_information.csv", header = TRUE)
```


```{r}
subinfo <-sinfo
subinfo <- subinfo %>%
  filter(!(workerid %in% blacklist))
```


```{r}
ggplot(subinfo, aes(x=race)) + geom_bar()
```

df with race encoded
```{r}
whitenames = list("Peter", "Brad", "Ethan", "Ian", "Cody", "Brett", "Paul", "Connor", "Jack", "Logan", "Roger", "Dylan", "Hunter", "Dustin","Ryan")
blacknames = list("Trevon", "Tyree", "Deion", "Marquis", "Jermaine", "Lamont", "Tyrone", "Deandre", "Tremayne", "Lamar", "Kareem", "Hakeem", "Jamal", "Rasheed", "Deshawn")
d  <-  mutate(d, stimRace = ifelse(first %in% whitenames, "White",
                                     ifelse( first %in% blacknames, "Black",
                                             "None")))
d <- d %>% select(-audio, -image, -slide_number, -slide_type, -Answer.condition)
d <- d %>% drop_na(trial_num)
```

reformatting some data?
```{r}
d$response <-  as.numeric(as.character(d$response))
```

creating important dataframes that are useful
```{r}
df.critical <- filter(d, type == "critical")
df.criticalW <- filter(df.critical, stimRace == "White")
df.criticalB <- filter(df.critical, stimRace == "Black")
```



starting to visualize some things…

looking at average responses by race/mean rating
```{r}
ERmeans <- aggregate(response ~ tag, d, mean)
ERmeans
```


start with average responses by race/mean rating by evoke or suppress criminality
```{r}
ERbyrace <- ggplot(ERmeans, aes(x = tag, y = response)) +
  geom_bar(stat = "identity", position =) +
  xlab("Race") +
  ylab("Mean Rating")
ERbyrace

raceER <- ggplot(d, aes(x=tag, y=mean(response), fill=stimRace)) +
  geom_col(stat = "identity", position = "dodge")
raceER
```

```{r}
nameMeans <- aggregate(response ~ name, df, mean)
nameMeans <- nameMeans[-1,]
nameMeans <- mutate(nameMeans, race = ifelse(name %in% whitenames, "White", "Black"))
nameMeans <-  arrange(nameMeans, race)

ggplot(nameMeans, aes(x = name, y = response, fill = race)) + 
  geom_bar(stat="identity") +
  theme(axis.text.x = element_text(angle = 80, hjust = 1))
```


trying more things
```{r}
evokedf <- filter(d, tag == "evoke")
repressdf <- filter(d, tag == "repress")

nameMeans1 <- aggregate(response ~ first, evokedf, mean)
nameMeans1 <- mutate(nameMeans1, race = ifelse(first %in% whitenames, "White", "Black"))
nameMeans1 <- arrange(nameMeans1, race)

nameMeans2 <- aggregate(response ~ first, repressdf, mean)
nameMeans2 <- mutate(nameMeans2, race = ifelse(first %in% whitenames, "White", "Black"))
nameMeans2 <- arrange(nameMeans2, race)


nM <- cbind(as.character(nameMeans1$first), nameMeans1$response, nameMeans2$response, nameMeans1$race)

nM1 <- tibble(as.character(nameMeans1$first),
              nameMeans1$response,
              nameMeans2$response,
              nameMeans1$race)
colnames(nM1) <- c("first", "evokeMean", "repressMean", "race")
```

visualize data by specific name
```{r}
nameDataE <- ggplot(nM1, aes(x=first, y=evokeMean, color = race)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 80, hjust = 1))

nameDataR <- ggplot(nM1, aes(x=first, y=repressMean, color = race)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 80, hjust = 1))

nameDataE
nameDataR
```


visualize by story data
```{r}
storyMeansE <- aggregate(response ~ story, evokedf, mean)
storyMeansE <- mutate(storyMeansE, race = ifelse(first %in% whitenames, "White", "Black"))

storyMeansR <- aggregate(response ~ story, repressdf, mean)
storyMeansR <- mutate(storyMeansR, race = ifelse(first %in% whitenames, "White", "Black"))

sM <- tibble(as.character(storyMeansR$story),
              storyMeansE$response,
              storyMeansR$response,
              storyMeansR$race)

colnames(sM) <- c("story", "evokeMean", "repressMean")

ggplot(sM, aes(x = story, y = evokeMean)) + geom_bar(stat="identity") +
  theme(axis.text.x = element_text(angle = 80, hjust = 1))
ggplot(sM, aes(x = story, y = repressMean)) + geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 80, hjust = 1))

ggplot(sM, aes(x=story, y = repressMean - evokeMean)) + geom_bar (stat = "identity") +
  theme(axis.text.x = element_text(angle = 80, hjust = 1))
```

Basic visualization plot
```{r}
bPlot <- ggplot(byStereotypeNameCategory, aes(x=nameCategory, y=Mean, fill = stereotype)) +
  # facet_wrap(~kind, scales = "free") +
  theme_bw() +
  geom_bar(stat="identity",position = "dodge") +
  theme(axis.text.x=element_text(angle=20,hjust=1,vjust=1)) +
  geom_errorbar(aes(ymin=YMin,ymax=YMax),size = 0.25,width= 0.025,position = dodge) +  
  scale_fill_manual("Item effect on\ncriminality stereotype", values = c("#00397A","#CCCCCC")) +
  labs(x = "Race", y = "Mean rating", fill = "Stereotype")  +
  scale_x_discrete(labels = c("White", "Black")) +
  ggtitle("Mean responses for White- and Black-normed names")

bPlot
```


#Looking at plots for the presentation and final paper
looking at variation in response by self-reported political beliefs
```{r}
#Faceting by political means
##adding political beliefs to dataframe
affiliationList <- rep(subinfo$affiliation, each = 24)
dfAffiliation <- cbind(df, affiliationList) 
dfAffiliation <- dfAffiliation %>% rename(affiliation = affiliationList)
dfAffiliation <- mutate(dfAffiliation, affiliationBin = ifelse(affiliation %in% 0:2, "Democrat",
                                                      ifelse(affiliation == 3, "Independent",
                                                      ifelse(affiliation %in% 4:6, "Republican",
                                                      "Unreported"))))

affiliationLabs <- c("No answer given", "Liberal Democrat\n(most liberal)", "Democratic-leaning Independent", "Moderate Democrat", "Independent\n(most central)", "Moderate Republican", "Republican-leaning Independent", "Conservative Republican\n(most conservative)")
affiliationLabsSimple <- c("No answer given", "Most liberal", "", "", "Most central", "", "", "Most conservative")
affiliationBinLabs <- c("Democrat","Independent","Republican", "Unreported")

##getting the numbers
byAffiliation <- dfAffiliation %>% 
  filter(type == "critical") %>%
  group_by(nameCategory, stereotype, affiliation) %>%
  summarize(Mean = mean(response), 
            CILow =ci.low(response),
            CIHigh =ci.high(response)) %>%
  ungroup() %>%
  mutate(YMin = Mean - CILow,
         YMax = Mean + CIHigh)

byAffiliationBin <- dfAffiliation %>% 
  filter(type == "critical") %>%
  group_by(nameCategory, stereotype, affiliationBin) %>%
  summarize(Mean = mean(response), 
            CILow =ci.low(response),
            CIHigh =ci.high(response)) %>%
  ungroup() %>%
  mutate(YMin = Mean - CILow,
         YMax = Mean + CIHigh)

byAffiliationStory <- dfAffiliation %>%
  filter(type == "critical") %>%
  group_by(nameCategory, stereotype, item, affiliation) %>%
  summarize(Mean = mean(response), 
            CILow =ci.low(response),
            CIHigh =ci.high(response)) %>%
  ungroup() %>%
  mutate(YMin = Mean - CILow,
         YMax = Mean + CIHigh)
```

```{r}
#Means plots collapsing across item
affiliationPlotByItem <- ggplot(byAffiliationStory, aes(x=factor(affiliation), y=Mean, fill = nameCategory)) +
  theme_bw() +
  #scale_fill_grey() +
  geom_bar(stat="identity",position = "dodge") +
  theme(axis.text.x=element_text(angle=70,hjust=1,vjust=1)) +
  ggtitle("Mean response on different items by subjects' self-reported political beliefs") +
  geom_errorbar(aes(ymin=YMin,ymax=YMax),size = 0.25,width= 0.025,position = dodge) +
  labs(x = "Political Affiliation", y = "Mean rating", fill = "Stereotype")  +
  scale_fill_manual("Race of\nnormed name", values = c("#00397A","#CCCCCC")) +
  facet_wrap(~item) +
  scale_x_discrete(labels = affiliationLabsSimple)
  #scale_x_discrete(labels = affiliationLabs)

affiliationPlotByItem
```

```{r}
#Means plots not collapsing across item
affiliationPlot <- ggplot(byAffiliation, aes(x=factor(affiliation), y=Mean, fill = nameCategory)) +
  facet_wrap(~stereotype, scales = "free") +
  theme_bw() +
  #scale_fill_grey() +
  geom_bar(stat="identity",position = "dodge") +
  theme(axis.text.x=element_text(angle=70,hjust=1,vjust=1)) +
  ggtitle("Mean response by subjects' self-reported political beliefs") +
  geom_errorbar(aes(ymin=YMin,ymax=YMax),size = 0.25,width= 0.025,position = dodge) +
  labs(x = "Political Affiliation", y = "Mean rating", fill = "Stereotype")  +
  scale_fill_manual("Race of\nnormed name", values = c("#00397A","#CCCCCC")) +
  scale_x_discrete(labels = affiliationLabsSimple) +
  ylim(0, 70)
  #scale_x_discrete(labels = affiliationLabs)

affiliationPlot
```

```{r}
#revised affiliation bin plot
byAffiliationBin$facet = factor(byAffiliationBin$affiliationBin, 
                                levels = c("Democrat", "Republican",
                                           "Independent","Unreported"))
#                               levels = c("Democrat (n=65)", "Republican (n=38)",
#                                           "Independent (n=21)","Unreported (n=1)"))


affiliationBinPlot <- ggplot(byAffiliationBin, aes(x=nameCategory, y=Mean, fill = stereotype)) +
  facet_wrap(~facet, scales = "free") +
  theme_bw() +
  geom_bar(stat="identity",position = "dodge") +
  theme(axis.text.x=element_text(angle=20,hjust=1,vjust=1)) +
  geom_errorbar(aes(ymin=YMin,ymax=YMax),size = 0.25,width= 0.025,position = dodge) +  
  scale_fill_manual("Item effect on\ncriminality stereotype",
                    values = c("#00397A","#CCCCCC")) +
  labs(x = "Stimulus Race", y = "Mean Response", fill = "Stereotype")  +
  scale_x_discrete(labels = c("White", "Black")) +
  ylim(0, 63) +
  ggtitle("Mean response by subjects' self-reported political beliefs")
affiliationBinPlot

```


subsetting off just the white speakers (who are vastly over-represented in the data, unsurprisingly) by facetting by reported race
```{r}
#Labels
raceLabs <- c("Unreported", "African American", "African American & White", "Asian", "Asian & African American", "Asian & White", "Latino/a/x", "Latino/a/x & White", "White")
names(raceLabs) <- c("", "AA", "AA, White", "Asian", "Asian, AA", "Asian, White", "Latino", "Latino, White", "White")

#Faceting by race for overall means
subRaceList <- rep(subinfo$race, each = 24)
dfSubRace <- cbind(df, subRaceList) 
dfSubRace <- dfSubRace %>% rename(subjectRace = subRaceList)

dfSubRace <- mutate(dfSubRace, raceBin = ifelse(subjectRace == "Asian", "Asian",
                                         ifelse(subjectRace == "Asian, White", "Asian Biracial",
                                         ifelse(subjectRace == "AA", "African American",
                                         ifelse(subjectRace %in% c("AA, White", "Asian, AA"),
                                                "African American Biracial",
                                         ifelse(subjectRace == "Latino", "Latino/a/x",
                                         ifelse(subjectRace == "Latino, White", "Latino/a/x Biracial",
                                         ifelse(subjectRace == "White", "White",
                                                "Unreported"))))))))

#New dataframe
bySubRace <- dfSubRace %>%
  filter(type == "critical") %>%
  group_by(nameCategory, stereotype, subjectRace, raceBin) %>%
  summarize(Mean = mean(response), 
            CILow =ci.low(response),
            CIHigh =ci.high(response)) %>%
  ungroup() %>%
  mutate(YMin = Mean - CILow,
         YMax = Mean + CIHigh)
```

```{r}
#Graphing
subRacePlot <- ggplot(bySubRace, aes(x=nameCategory, y=Mean, fill = stereotype)) +
  #facet_wrap(~subjectRace, scale = "free", labeller = labeller(subjectRace = raceLabs)) +
  facet_wrap(~subjectRace, scale = "free") +
  theme_bw() +
  #scale_fill_grey() +
  scale_fill_manual("Item effect on\ncriminality\nstereotype", values = c("#00397A","#CCCCCC")) +
  geom_bar(stat="identity",position = "dodge") +
  #theme(axis.text.x=element_text(angle=0,hjust=1,vjust=1)) +
  geom_errorbar(aes(ymin=YMin,ymax=YMax),size = 0.25,width= 0.025,position = dodge) +
  labs(x = "Stimulus Race", y = "Mean rating", fill = "Subject Race")+
  ggtitle("Mean responses by subject's self-reported race")

subRacePlot
```

```{r}
binFacets <- c("African American (n=8)", "African American Biracial (n=3)",
                           "Asian (n=7)", "Asian Biracial (n=1)", "Latino/a/x (n=8)",
                           "Latino/a/x Biracial (n=4)", "Unreported (n=1)", "White (n=93)")
```

```{r}
#binned plot
subRaceBinPlot <- ggplot(bySubRace, aes(x=nameCategory, y=Mean, fill = stereotype)) +
  facet_wrap(~ raceBin, labeller = labeller(binFacets)) +
  #facet_wrap(~raceBin, scale = "free") +
  theme_bw() +
  #scale_fill_grey() +
  scale_fill_manual("Item effect on\ncriminality\nstereotype", values = c("#00397A","#CCCCCC")) +
  geom_bar(stat="identity",position = "dodge") +
  #theme(axis.text.x=element_text(angle=0,hjust=1,vjust=1)) +
  geom_errorbar(aes(ymin=YMin,ymax=YMax),size = 0.25,width= 0.025,position = dodge) +
  #labs(x = "Stimulus Race", y = "Mean rating", fill = "Subject Race")+
  labs(x = "Stimulus Race", y = "Mean rating")+
  ylim(0, 85) +
  ggtitle("Mean responses by subject's self-reported race")

subRaceBinPlot
```


looking at the by-worker patterns of response
```{r}
byworkerPlot <- ggplot(byStereotypeNameCategoryWorker, aes(x=nameCategory, y=Mean, fill = stereotype)) +
  facet_wrap(~workerid) +
  theme_bw() +
  #scale_fill_grey() +
  scale_fill_manual("Item effect on\ncriminality\nstereotype", values = c("#00397A","#CCCCCC")) +
  geom_bar(stat="identity",position = "dodge") +
  theme(axis.text.x=element_text(angle=20,hjust=1,vjust=1)) +
  geom_errorbar(aes(ymin=YMin,ymax=YMax),size = 0.25,width= 0.025,position = dodge) +  
  labs(x = "Race", y = "Mean rating", fill = "Stereotype")+
  ggtitle("Mean response by individual participant")

byworkerPlot
```


visualizations of the by-item variability
```{r}
byStereotypeNameCategoryStory <- df %>%
  filter(type == "critical") %>%
  group_by(nameCategory, stereotype, item) %>%
  summarize(Mean = mean(response), 
            CILow =ci.low(response),
            CIHigh =ci.high(response)) %>%
  ungroup() %>%
  mutate(YMin = Mean - CILow,
         YMax = Mean + CIHigh)
```

```{r}
#Repress condition
itemMeansR <- ggplot(filter(byStereotypeNameCategoryStory, stereotype == "repress"), aes(x=item, y=Mean, fill = nameCategory)) +
  geom_hline(yintercept = mean(filter(byStereotypeNameCategoryStory, stereotype == "repress")$Mean), color="gray") +
  theme_bw() +
  #scale_fill_grey() +
  scale_fill_manual("Race of\nnormed name", values = c("#00397A","#CCCCCC")) +
  geom_bar(stat="identity",position = "dodge") +
  theme(axis.text.x=element_text(angle=70,hjust=1,vjust=1)) +
  geom_errorbar(aes(ymin=YMin,ymax=YMax),size = 0.25,width= 0.025,position = dodge) +  
  ylim(0, 85) +
  labs(x = "Presented story", y = "Mean rating", fill = "Race")+
  ggtitle("Average response by items repressing criminality stereotype")

itemMeansR
```

```{r}
#Evoke condition
itemMeansE <- ggplot(filter(byStereotypeNameCategoryStory, stereotype == "evoke"), aes(x=item, y=Mean, fill = nameCategory)) +
  geom_hline(yintercept = mean(filter(byStereotypeNameCategoryStory, stereotype == "evoke")$Mean), color="gray") +
  theme_bw() +
  #scale_fill_grey() +
  scale_fill_manual("Race of\nnormed name", values = c("#00397A","#CCCCCC")) +
  geom_bar(stat="identity",position = "dodge") +
  theme(axis.text.x=element_text(angle=70,hjust=1,vjust=1)) +
  geom_errorbar(aes(ymin=YMin,ymax=YMax),size = 0.25,width= 0.025,position = dodge) + 
  ylim(0, 85) +
  labs(x = "Presented story", y = "Mean rating", fill = "Race")+
  ggtitle("Average response by items evoking criminality stereotype")

itemMeansE

```


ordering effects: whether/how responses differ between the first and second half of the study
overall plot of means and facet by first vs second half, that should be enough to see visually whether there's any order effect
```{r}
dfOrder <- mutate(df, orderHalf = ifelse(trial_num < 12, "First Half", "Second Half"))

byStereotypeNameCategoryOrder <- dfOrder %>%
  filter(type == "critical") %>%
  group_by(nameCategory, stereotype, orderHalf) %>%
  summarize(Mean = mean(response), 
            CILow =ci.low(response),
            CIHigh =ci.high(response)) %>%
  ungroup() %>%
  mutate(YMin = Mean - CILow,
         YMax = Mean + CIHigh)

ordering <- ggplot(byStereotypeNameCategoryOrder, aes(x=nameCategory, y=Mean, fill = stereotype)) +
  # facet_wrap(~kind, scales = "free") +
  theme_bw() +
  #scale_fill_grey() +
  scale_fill_manual("Race of\nnormed name", values = c("#00397A","#CCCCCC")) +
  geom_bar(stat="identity",position = "dodge") +
  facet_wrap(~orderHalf) +
  theme(axis.text.x=element_text(angle=20,hjust=1,vjust=1)) +
  geom_errorbar(aes(ymin=YMin,ymax=YMax),size = 0.25,width= 0.025,position = dodge) +  
  labs(x = "Race", y = "Mean rating", fill = "Stereotype") +
  ggtitle("Response means by order in the experiment")

ordering
```

histograms of workers
```{r}
#subject race
raceHist <- ggplot(subinfo, aes(x = race)) +
  geom_histogram(stat = "count", color = "#CCCCCC", fill = "#CCCCCC") +
  scale_x_discrete(labels = raceLabs) +
  theme_bw() +
  labs(x = "Race", y = "Frequency") +
  theme(axis.text.x=element_text(angle=30,hjust=1,vjust=1)) +
  ggtitle("Frequency of subjects' self-reported race")

sinfoRaceBin <- mutate(subinfo, raceBin = ifelse(race == "Asian", "Asian",
                                         ifelse(race == "Asian, White", "Asian Biracial",
                                         ifelse(race == "AA", "African American",
                                         ifelse(race %in% c("AA, White", "Asian, AA"),
                                                "African American Biracial",
                                         ifelse(race == "Latino", "Latino/a/x",
                                         ifelse(race == "Latino, White", "Latino/a/x Biracial",
                                         ifelse(race == "White", "White",
                                                "Unreported"))))))))
```

```{r}
raceBinHist <- ggplot(sinfoRaceBin, aes(x = raceBin)) +
  geom_histogram(stat = "count", color = "#CCCCCC", fill = "#CCCCCC") +
  #scale_x_discrete(labels = raceLabs) +
  theme_bw() +
  labs(x = "Race", y = "Frequency") +
  theme(axis.text.x=element_text(angle=30,hjust=1,vjust=1)) +
  ggtitle("Frequency of subjects' self-reported race")

raceBinHist
```

```{r}
#subject political affiliation
politicalHist <- ggplot(subinfo, aes(x = factor(affiliation))) +
  geom_histogram(stat = "count", color = "#CCCCCC", fill = "#CCCCCC") +
  scale_x_discrete(labels = affiliationLabs) +
  theme_bw() +
  labs(x = "Political Affiliation", y = "Frequency") +
  theme(axis.text.x=element_text(angle=30,hjust=1,vjust=1)) +
  ggtitle("Frequency of subjects' self-reported political affiliation")

politicalHist
```

```{r}
dfPoliticalBin <- mutate(subinfo, affiliationBin = ifelse(affiliation %in% 0:2,
                                                                   "Democrat",
                                                                   ifelse(affiliation == 3,
                                                                    "Independent",
                                                                    ifelse(affiliation %in%
                                                                      4:6, "Republican",
                                                                      "Unreported"))))

politicalBinHist <- ggplot(mutate(subinfo, affiliationBin = ifelse(affiliation %in% 0:2,
                                                                   "Democrat",
                                                                   ifelse(affiliation == 3,
                                                                    "Independent",
                                                                    ifelse(affiliation %in%
                                                                      4:6, "Republican",
                                                                      "Unreported")))), 
                           aes(x = factor(affiliationBin))) +
  geom_histogram(stat = "count", color = c("#00397A", "#CCCCCC", "#B51700", "#000000"), 
                 fill = c("#00397A", "#CCCCCC", "#B51700", "#000000")) +
  scale_x_discrete(labels = affiliationBinLabs) +
  theme_bw() +
  labs(x = "Political Affiliation", y = "Frequency") +
  theme(axis.text.x=element_text(angle=30,hjust=1,vjust=1)) +
  ggtitle("Frequency of subjects' self-reported political affiliation")

politicalBinHist
```
